Wireless Automobile Detection, License Plate Processing, and Data Availability Network

### 2014-05-01 paper/paper.tex
Wrote about results and future work, image

\begin{abstract}
{\color{red}\textbf{NOT COMPLETE}}
This paper describes a parking lot monitoring system to be used at the Colorado School of Mines (CSM). The system will help campus members enjoy a better parking experience by spending less time trying to find open lots. This is a semester project for the Computer Science graduate course \emph{Wireless Sensor Networks}, taught by Dr. Tracy Camp at CSM. The work was an extension of a project started by campus computing group ACMx, which uses a wireless sensor network to track the number of cars in specific parking lots. The system will extend the ACMx project by adding components to the sensor network that record the license plate numbers of cars as they enter and exit lots. The information gathered by the sensor network will be displayed on a website that will allow users to survey the state of the lots before arriving, allowing them to make informed parking decisions. The website will also be the result of collaboration between the authors of this proposal and the ACMx group. The proposal provides some background information, describes the proposed system and its components, and discusses implementation strategies. The goal of the project is a real-world deployment on the CSM campus by May 2014.
\end{abstract}

\section{Introduction}
The parking lots at the Colorado School of Mines (CSM) can be a source of frustration for students and faculty members. Because space on campus is limited, the lots are often unable to accommodate everyone who needs to use them. Students need to choose a lot before they go to class; however, knowing which lots have free spaces usually requires driving through the lots looking for a space. The extra time spent searching one, or many, lots is frustrating, and can make students or faculty late for class. If campus members could discover which lots had available spaces prior to arrival, much time and frustration could be saved. This project involved the design and creation of a working parking lot monitoring system that will be the first step towards a better parking experience on the CSM campus. The system keeps track of both the number of vehicles within a parking lot and the license plate numbers of those vehicles. The information is made available online to system administrators and CSM campus members. There are two main parts to this project:

@@

The idea behind the first part is that a student can get on their smartphone, tablet, or laptop before heading to class, see a nicely-formatted list/map of the lots on campus, and decide where to park based on how full each lot is. Administrators can monitor the information and make sure it is correct, updating the web application as needed with new statistics, lists, and maps as more lots are added to the system. Before beginning this project, the Association for Computing Machinery (ACMx) student organization at CSM had already started working on a system for monitoring lot car counts and capacities (see section 1.2\textbf{\color{red} FIX SECTION NUMBER}). We collaborated with the ACMx group, integrating our new systems and work with theirs to design and test our parking lot monitoring system.

The second part of the project is motivated by a desire to test the feasibility of using the small Raspberry Pi Linux computer as a platform for capturing and processing images of license plates. Because the Raspberry Pi is inexpensive (\$25-\$35 depending on the model), it could be useful in a wide variety of applications that require accurate license plate recognition. Currently, our web application simply lists the license plate numbers for site administrators to view, but we envision that this monitoring system may eventually be used by the campus parking department to detect when vehicles have entered lots without parking passes. The system itself will consist of several subsystems, which are described in detail in section 2.

@@

\section{Project Overview and Methods}

\begin{figure}
\begin{center}
\includegraphics[width=4.5in]{architecture}
\end{center}
\caption{Overall System Architecture}
\label{fig:system}
\end{figure}

This project is partially an extension of the ACMx SmartLots project. The system consists of several components, each of which will is either an extension of a current SmartLots system or an entirely new subsystem that has been integrated into the existing system. Figure \ref{fig:system} shows the overall architecture of our system. The system relies on three main components:
\begin{enumerate}
\item Detection Subsystem,
\item Central Base Station, and
\item Server.
\end{enumerate}

@@

\section{Results}
\subsection{ALPR Results}
Three data sets were run through each ALPR system. In each case, Intrada ALPR was able to recognize at least twice as many plates as OpenALPR.

\begin{table}[h!]
\begin{center}
\begin{tabular}{| l || l | l | l |}
\hline
Data Set & Correct Recognition & Total Images & Percent Correct \\ \hline
Unaltered images & 12 & 41 & 29.27 \% \\ \hline
Cropped images & 4 & 52 & 7.69 \% \\ \hline
Compressed images & 4 & 27 & 14.81 \% \\ \hline
\end{tabular}
\end{center}
\caption{OpenALPR Results}
\end{table}

\begin{table}[h!]
\begin{center}
\begin{tabular}{| l || l | l | l |}
\hline
Data Set & Correct Recognition & Total Images & Percent Correct \\ \hline
Unaltered images & 27 & 41 & 65.85 \% \\ \hline
Cropped images & 12 & 52 & 23.08 \% \\ \hline
Compressed images & 8 & 27 & 29.63 \% \\ \hline
\end{tabular}
\end{center}
\caption{Intrada ALPR Results}
\end{table}

\section{Future Research}

>>>>

\begin{abstract}
{\color{red}\textbf{NOT COMPLETE}}
This proposal describes a parking lot monitoring system to be used at the Colorado School of Mines (CSM). The system will help campus members enjoy a better parking experience by spending less time trying to find open lots. This is a semester project for the Computer Science graduate course \emph{Wireless Sensor Networks}, taught by Dr. Tracy Camp at CSM. The proposed work will be an extension of a project started by campus computing group ACMx, which uses a wireless sensor network to track the number of cars in specific parking lots. The system will extend the ACMx project by adding components to the sensor network that record the license plate numbers of cars as they enter and exit lots. The information gathered by the sensor network will be displayed on a website that will allow users to survey the state of the lots before arriving, allowing them to make informed parking decisions. The website will also be the result of collaboration between the authors of this proposal and the ACMx group. The proposal provides some background information, describes the proposed system and its components, and discusses implementation strategies. The goal of the project is a real-world deployment on the CSM campus by May 2014.
\end{abstract}

\section{Introduction}
The parking lots at the Colorado School of Mines (CSM) can be a source of frustration for students and faculty members. Because space on campus is limited, the lots are often unable to accommodate everyone who needs to use them. Students need to choose a lot before they go to class; however, knowing which lots have free spaces usually requires driving through the lots looking for a space. The extra time spent searching one or many lots is frustrating, and can make students or faculty late for class. If campus members could discover which lots had available spaces before they arrived, much time and frustration could be saved. The goal of this project is to design and create a working parking lot monitoring system that will be the first step towards a better parking experience on the CSM campus. The system will keep track of both the number of vehicles in a parking lot and the license plate numbers of those vehicles. The information will be available online to system administrators and CSM campus members. There are two main parts to this project:

@@

The idea behind the first part is that a student can get on their smartphone, tablet, or laptop before heading to class, see a nicely-formatted list/map of the lots on campus, and decide where to park based on how full each lot is. Administrators will monitor the information and make sure it is correct, updating the web application as needed with new statistics, lists, and maps as more lots are added to the system. Before beginning this project, the Association for Computing Machinery (ACMx) student organization at CSM had already started working on a system for monitoring lot car counts and capacities (see section 1.2). We collaborated closely with the ACMx group, integrating our work with theirs to design and test our parking lot monitoring system.

The second part of the project is motivated by a desire to test the feasibility of using the small Raspberry Pi Linux computer as a platform for capturing and processing images of license plates. Because the Raspberry Pi is inexpensive (\$25-\$35 depending on the model), it could be useful in a wide variety of applications that require accurate license plate recognition. For now our web application will simply list the license plate numbers for site administrators to view, but we envision that this monitoring system may eventually be used by the campus parking department to detect when vehicles have entered lots without parking passes.

The goal of this project is to work with the ACMx group to create and deploy the system in two parking lots on campus by the end of the semester. See section 2.7 for information about the deployment. The system itself will consist of several subsystems, which are described in detail in section 2.

@@

\section{Project Overview and Methods}
\begin{figure}
\begin{center}
\includegraphics[width=4.5in]{architecture}
\end{center}
\caption{Overall System Architecture}
\label{fig:system}
\end{figure}
This project is partially an extension of the ACMx SmartLots project. The system consists of several components, each of which will is either an extension of a current SmartLots system or an entirely new subsystem that has been integrated into the existing system. Figure \ref{fig:system} shows the overall architecture of our system. The system relies on three main components:
\begin{enumerate}
\item Detection Subsystem,
\item Central Base Station, and
\item Server.
\end{enumerate}

@@

\section{Results}
\subsection{ALPR Results}
\begin{table}[h!]
\begin{center}
\begin{tabular}{| l || l | l | l |}
\hline
Data Set & Correct Recognition & Total Images & Percent Correct \\ \hline
Unaltered images & 20 & 52 & 38.46 \% \\ \hline
Cropped images & 4 & 52 & 7.69 \% \\ \hline
Compressed images & 4 & 27 & 14.81 \% \\ \hline
\end{tabular}
\end{center}
\caption{OpenALPR Results}
\label{tab:openalprresults}
\end{table}

\begin{table}[h!]
\begin{center}
\begin{tabular}{| l || l | l | l |}
\hline
Data Set & Correct Recognition & Total Images & Percent Correct \\ \hline
Unaltered images & 34 & 52 & 65.39 \% \\ \hline
Cropped images & 12 & 52 & 23.08 \% \\ \hline
Compressed images & 8 & 27 & 29.63 \% \\ \hline
\end{tabular}
\end{center}
\caption{Intrada ALPR Results}
\label{tab:intradaresults}
\end{table}
Three data sets were run through each ALPR system. As seen in Table \ref{tab:openalprresults} and Table \ref{tab:intradaresults}, Intrada ALPR outperformed OpenALPR in each case in terms of accuracy. In terms of speed, OpenALPR always outperformed Intrada ALPR by at least a factor of three. In some cases, OpenALPR was better by a factor of 110.

\begin{figure}
\begin{center}
\includegraphics[width=4.5in]{unexpectedrecognition}
\end{center}
\caption{The foreground license plate goes unrecognized in favor of a smaller plate in the background}
\label{fig:unexpectedrecognition}
\end{figure}

These results were calculated by determining whether or not the ALPR software correctly identified \emph{any} license plate in the photo. It is the case that, for certain out of state plates or nonstandard plates (e.g. Disabled Veteran), the ALPR programs will not identify the primary, or largest, license plate in the photo. However, both programs were successful in recognizing a much smaller, but standard, license plate. Figure \ref{fig:unexpectedrecognition} shows an example image where a New Mexico plate is ignored in favor of a smaller plate in the background.
\section{Future Research}

\subsection{ALPR Training}
The accuracy rate for our ALPR software is not as high as we had originally hoped it to be. OpenALPR offers a separate program, 'train-ocr' to help build definition files for certain geographic regions. Because it initially came seeded with training data for the entire continent, the recognition of Colorado plates may be improved if we create a training file for only Colorado.

Intrada ALPR cannot be configured by the end user (our team). We can, however, send sample images from our deployment to Q-Free, where our files can be used to train our account for better recognition.

Because we ended up using two ALPR systems instead of one, it would be beneficial to utilize both to provide a confidence level for our project. Having validation from both ALPRs would likely mean the characters that were extracted were correct. Extractions that were not correct could be flagged in the database for human intervention.

A proposed long-term goal for this project was to integrate license recognition with CSM's parking database to identify users who do not have valid parking permits. With the work that has been completed, we are now at a place where we could begin coding an interface to pass license data to Parking and Facilities.
